# Nonparametric Test

## One-sample location

### Sign Test

Suppose we have a sample of $X_1,X_2,...,X_n$ an arbitary **continuous** distribution. We use the median $m$ as denote their location parameter(In normal population, we use mean instead). Therefore, $P(X_i>m) = 0.5$ and $P(X_i<m) = 0.5$. We want to test
$$
H_0: m =m_0
$$
Let 
\begin{align}
Y_i &= \begin{cases} 1, X_i-m_0>0 \\
0, X_i-m_0<0
\end{cases}
B &= \sum_{i=1}^n Y_i
\end{align}
From the derivation of $B$, we can see $B\sim Bin(n,0.5)$, then $E(B) = n/2$ and $Var(B) = n/4$. Let $b^*$ be the observed value of $B$. The p-value of testing $H_0:m=m_0$ is 
$$
p = 
\left\{\begin{array}{ll}
P\left(B \geq b^{*}\right)=\sum_{b=b^{*}}^{n}\left(\begin{array}{l}
n \\
b
\end{array}\right)\left(\frac{1}{2}\right)^{n} & \text { if } H_{1}: m>m_{0} \\
P\left(B \leq b^{*}\right)=\sum_{b=0}^{b^{*}}\left(\begin{array}{l}
n \\
b
\end{array}\right)\left(\frac{1}{2}\right)^{n} & \text { if } H_{1}: m<m_{0} \\
2 \min \left\{\sum_{b=b^{*}}^{n}\left(\begin{array}{l}
n \\
b
\end{array}\right)\left(\frac{1}{2}\right)^{n}, \sum_{b=0}^{b^{*}}\left(\begin{array}{l}
n \\
b
\end{array}\right)\left(\frac{1}{2}\right)^{n}\right\} & \text { if } H_{1}: m \neq m_{0}
\end{array}\right.
$$

### Wilcoxon Signed Rank Test

Suppose we have a sample of $X_1,X_2,...,X_n$ an arbitary **continuous and symmetric**  distribution with median $m$. We want to test 
$$
H_0:m=m_0
$$
We let $Y_i = X_i-m_0$, sort $|Y_i|$ in ascending order and denote the corresponding rank $R_i$, let 
\begin{equation}
S_i = \begin{cases} 1, Y_i>0,\\
-1, Y_i < 0
\end{cases}
\end{equation}
$$
W = \sum_{i=1}^n R_iS_i
$$
with $E(W) = 0$ and $Var(W) = \frac{n(n+1)(2n+1)}{6}$.

## Example 

<div style="font-size: 30px">
```{r}
x <- c(5.625,5.665,5.697,5.837,5.863,5.870,5.878,5.884,5.908,
5.967,6.019,6.020,6.029,6.032,6.037,6.045,6.049,6.387, 
6.050,6.079,6.116,6.159,6.186,6.199,6.307)

n <- length(x)
B <- sum(x > 5.9) # number of X_i greater than m_0 = 5.9
p <- 1 - pbinom(B-1, n, 0.5)
cat("p-value calculated by hand: ", p)
binom.test(B, n, alternative = "greater")
```

```{r}
m0 <- 5.9
y <- x - m0
y_abs <- abs(y)
R <- rank(y_abs)
S <- as.integer(y > 0) - as.integer(y < 0)
rs1 <- R * S

dt <- data.frame(x, y, y_abs, R, S, rs1)
names(dt) <- c("X","Y","|y|","Rank of |y|","sign","sign x Rank")
knitr::kable(dt,booktabs = TRUE)

W <- sum(rs1)
rs2 <- R * as.integer(y>0)
p_wil <- 1 - pnorm((W - 1) / sqrt(n * (n + 1) * (2 * n + 1) / 6))
cat("p-value calculated by hand", p_wil)

wilcox.test(y, exact = FALSE, alternative = "greater")
```

```{r}
D <- x - m0
d <- mean(D)
sd <- sd(x)
test <- d / (sd / sqrt(n))
cat("d:", d, "sd: ", sd, "test T: ", test, "\n")
cat("p-value calculated by hand: ", 1 - pt(test, 24))
t.test(x - m0, alternative = "greater")
```
</div>